### 2023-06-01 paper/adv_pcg_convergence.tex
move adv pcg and geigs plots to separate files

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamart190516}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\input{localpsf_shared}

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels},
  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
}
\fi

% The next statement enables references to information in the
% supplement. See the xr-hyperref package for details.

\externaldocument{localpsf_supplement}

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

% CITATIONS TO DO
%
% TASKS:
% 1) Incorporate Ronald Kriemann's suggested citations: (Nick)
%    - Steffen Börm, Lars Grasedyck and Wolfgang Hackbusch:
%      "Introduction to Hierarchical Matrices with Applications",
%       Engineering Analysis with Boundary Elements 27(5), pp. 405--422, 2003. (DONE)
%    - Lars Grasedyck, Ronald Kriemann and Sabine Le Borne:
%      "Parallel blackbox H-LU preconditioning for elliptic boundary value problems",
%       Computing and Visualization in Sciences, Vol. 11, pp 273--291, 2007. (DONE)
%    - Ronald Kriemann:
%      "H-LU Factorization on Many-Core Systems",
%      Computing and Visualization in Science, 16, pp. 105-117, 2015. (DONE)
%
% 2) Grants
%    - Nick and Omar's grants (Omar)
%    - Tucker and Noemi's grants (DONE)
%
% 3) Order citations in ascending order within brackets (Nick)
%    e.g., [1,2,3] = good
%          [2,3,1] = bad
%
% 4) Find and cite relevant papers from Noemis's group
%    - Tucker's technical report (Tucker)
%
% 5) Find and cite relevant papers from Omar's group
%    - Toby's papers (Tucker)
%    - Noemi's 2012 stokes paper (Tucker)
%    - Omar and Karen's recent review paper (DONE)
%
% 6) Look through Nick's previous PC paper for missed citations (DONE)

% For citations we need:
%    a) bibtex info
%    b) exactly where in the paper to add the ref
%    c) any words or sentences to add with the ref

\newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{To do: #1\:}}}

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
%	, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). 
%	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators. 
	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel. We compute a large number of impulse responses by applying the operator to a small number Dirac combs associated with ``batches'' of point sources. By solving an ellipsoid packing problem, we compute as many impulse responses as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions. The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance. The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator. Further matrix computations are performed with H-matrix methods. We apply the method to approximate the Hessians in an ice sheet flow PDE constrained inverse problem, and an advection-diffusion PDE constrained inverse problem. Numerical results demonstrate that the proposed method outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
\end{abstract}

% REQUIRED
\begin{keywords}
 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response, moment methods
\end{keywords}

% REQUIRED
\begin{AMS}
 	35R30, 41A05, 41A35, 41A36, 42A85, 47A52, 47A58, 49K20, 65D12, 65F08, 65F10, 65N20, 65N21, 86A22, 86A40
\end{AMS}

\section{Introduction}
\label{sec:intro}

We present an efficient \emph{matrix-free} point spread function (PSF) method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels. Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$. We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AlgerEtAl19,IsaacEtAl15}. Such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}. Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$. }, $\Aop^T$, to arbitrary functions,
\begin{equation}
\label{eq:eval_maps}
u \mapsto\Aop u \qquad \text{and} \qquad w \mapsto \Aop^T w,
\end{equation} 
via a black box computational procedure, 
but we cannot easily access entries of $\Aop$'s integral kernel. Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or performing other costly computations.

\begin{figure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.33]{impulse_batch1.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.33]{impulse_batch2.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.33]{impulse_batch3.png}
	\end{subfigure}
	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the Stokes inverse problem data misfit Gauss-Newton Hessian (Section~\ref{sec:numerical_results}). 
%	Brightness scale differs between the subfigures. 
	Black stars are point source locations. Dashed gray ellipses are estimated impulse response support ellipsoids. The large circle is $\partial \Omega$. 
	}
	\label{fig:batches_intro}
\end{figure}

%Existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, which we cannot do. 
We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applies. The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$. We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure~\ref{fig:batches_intro}). Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel. 
%This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods. 
Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} we compute them. 
In resolution analysis in seismic imaging, 
$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}. Here, we estimate the support of the $\phi_x$ by applying $\Aop^T$ to polynomial functions instead of random noise functions (see Section~\ref{sec:intromoments}). Our method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.

Our method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights. PC and PSF methods have a long history dating back several decades. We note the following papers (among many others), \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources. For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}. While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}). 
%Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulses. We avoid this issue by temporarily excluding impulses from the approximation procedure whenever an attempt is made to evaluate the impulse at a location where it is undefined.

%\begin{figure}
%	\begin{center}
%		\includegraphics[scale=0.4]{bct1.eps}
%	\end{center}
%	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
%	}
%	\label{fig:hmatrix_intro}
%\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
	\end{center}
	\caption{$\Aop$ is locally mean displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$. Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
	\label{fig:mean_displacement_invariance}
\end{figure}

The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to approximate discretized versions of $\Aop$ using the full arsenal of tools for matrix approximation that rely on fast access to matrix entries.
%one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$. H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank. 
%(see Figure~\ref{fig:hmatrix_intro} and Appendix~\ref{app:h_matrix}).
H-matrix methods permit us to perform matrix-vector products cheaply, and perform other useful linear algebra operations that cannot be done easily using the original operator. These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion. 
%These H-matrix methods are fast and scale well to large problems. 
The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix scales 
%as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$. That is, the cost is \emph{nearly-linear}
\emph{nearly linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$). The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}. These operations can be used, for example, to solve Newton linear systems in PDE constrained optimization, sample from ill-conditioned posterior distributions in Bayesian inverse problems, and construct high rank Gaussian surrogate models that can be used for uncertainty quantification. 
%Classical H-matrix construction methods 
%(described in Appendix~\ref{sec:H_matrix_construction}) 
%require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.

\section{Why we need high rank Hessian approximations}
\label{sec:hessian}

In distributed parameter inverse problems governed by PDEs, one seeks to infer an unknown spatially varying parameter field from limited observations of a state variable that depends on the parameter implicitly through the solution of a PDE. 
%Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,Vogel02},
%Guy Chavent, Nonlinear Least Squares for Inverse Problems, Springer, 2009.
or a Bayesian probabilistic framework~\cite{KaipioSomersalo05,Stuart10a,Tarantola05}. In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization~\cite{EnglHankeNeubauer96}. In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior. The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations. For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.

Hessian approximations and preconditioners are highly desirable because the Hessian is central to efficient solution of inverse problems in both the deterministic and Bayesian settings. When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration. Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient~\cite{Saad03}. In the Bayesian setting, the inverse of the Hessian is the covariance of a local Gaussian approximation of the posterior. Such a Gaussian distribution can be used as a direct approximations of the posterior, or as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior. For instance, see~\cite{KimEtAl21,PetraEtAl14} and the references therein.

%\begin{figure}
%	\begin{subfigure}{0.5\textwidth}
%		\includegraphics[scale=0.23]{informed_uninformed_modes_quadratic.pdf}
%		\caption{}
%	\end{subfigure}
%	\begin{subfigure}{0.48\textwidth}
%		\includegraphics[scale=0.1]{misfit_and_reg_hessian_eigenvalues.pdf}
%		\caption{}
%	\end{subfigure}
%	\caption{(a) The Hessian eigenstructure characterizes the directional scalings of the best local quadratic approximation to the objective function in an inverse problem. The larger an eigenvalue of the Hessian is, the more closely spaced the contours of the local quadratic approximation are in the corresponding eigenvector direction. (b) As the data in an inverse problem becomes more informative about the parameter, the eigenvalues of the data misfit Hessian increase. As the regularization is weakened, the eigenvalues of the regularization Hessian decrease. These illustrations are taken from~\cite[Chapter 1]{Alger19}, with minor modifications.}
%	\label{fig:hessian_eigenstructure}
%\end{figure}

Owing to the implicit dependence of predicted observations on the parameter, entries of the Hessian are not easily accessible. Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21,Gunzburger03}. The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g.,~\cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}. Krylov methods such as Lanczos or randomized methods~\cite{Cheng05,HalkoMartinssonTropp11} are typically used to construct these low rank approximations by applying the Hessian to vectors. Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is proportional to the rank of the low rank approximation. Low rank approximation methods are justified by arguing that the numerical rank of the data misfit Hessian is insensitive to the dimension of the discretized parameter. This means that the required number of PDE solves remains the same as the mesh used to discretize the parameter is refined.
However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods~\cite{AmbartsumyanEtAl20,BuiEtAl13,IsaacEtAl15}.

Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems~\cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter. The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
%(see Figure~\ref{fig:hessian_eigenstructure})
so more informative data leads to larger eigenvalues and therefore a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}. Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data. The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization~\cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}. Hence, low rank approximation methods suffer from a predicament---if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods. 
%The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost)~\cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}. 
High rank Hessian approximation methods are needed.

Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation~\cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}. 
These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,HartlandEtAl23}. 
Although the required number of matrix-vector products grows nearly linearly in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, the required number of matrix-vector products is still large (e.g., hundreds to thousands).
%These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks. But despite asymptotic scalability, from a practical perspective the required number of matrix-vector products is large (e.g., hundreds to thousands). 
%Improving the peeling process to reduce its cost is currently an area of active research~\cite{PEELINGRESEARCH}.
%This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
%Matrix-free H-matrix construction methods are currently an area of active research (see, e.g.,~\cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future. 
In this paper, we also form an H-matrix Hessian approximation. However, to reduce the required number of operator actions, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques. Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel). However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and our method is targeted at approximating these Hessians.

%Classical methods for approximating the Hessian are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the inverse of the prior term in the Hessian. However,  More generally, the eigenvalues of the Hessian characterize how informative the data are about the component of the parameter in the corresponding parameter direction. The more informed the component of the parameter, the larger

%In the deterministic setting, the Hessian of the objective function is central to efficient solution of the optimization problem, and in the

% given the observed data is  one explores the posterior probability distribution for the parameter

%Our motivation is high rank Hessian approximation in distributed parameter inverse problems governed by PDEs. 
%The Hessian is an operator that is central to efficient methods for solving such inverse problems and is only available matrix-free; we will discuss Hessians in inverse problems governed by PDEs in greater detail in Section~\ref{sec:hessian}.

%Introduce inverse problem
%
%Role of Hessian
%
%Structure of Hessian that we want to exploit
%
%Low rank
%
%Need for High rank
%
%Low rank approximation of the Hessian:
%\begin{itemize} 
%	\item Hessian eigenstructure is determined by data informativeness. Bigger eigenvalue => more informed component in eigenvector direction. 
%	\item informative data => high rank hessian
%	\item get high rank approximation by exploiting different structure. Locality, non-negativity of kernel
%	\item Large rank in absolute terms, even if low rank in relative terms
%\end{itemize}

\section{Preliminaries}

Let $\Omega \subset \mathbb{R}^\gdim$ be a bounded domain (typically $\gdim=1$, $2$, or $3$). We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
\begin{equation}
\label{eq:kernel_representation}
(\Aop u)(w) := \int_\Omega \int_\Omega w(y) \Aker(y,x) u(x) dx dy.
\end{equation}
Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$.
The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible. In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section~\ref{sec:conditions_2}), and we detail finite element discretization (Section~\ref{sec:finite_element_kernel}).

\subsection{Distributions and impulse responses}
\label{sec:impulse_response_2}

The action of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. 
%We derive the extension as follows. 
Let $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
\begin{subequations}
	\label{eq:extension_to_distributions}
	\begin{align}
		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy 
		= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
	\end{align}
\end{subequations}
where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution. In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in~\eqref{eq:extension_to_distributions} is not valid. However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of~\eqref{eq:action_on_distribution} is well defined. Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}. We denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.

Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is defined by $\delta_x(w) = w(x)$ for all $w\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
\begin{equation}
\label{eq:impulse_response_delta_action}
\impulseresponse_x := \left( \Aop \delta_x^* \right)^*,
\end{equation}
that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional. Using~\eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function $\impulseresponse_x(y) = \Aker(y, x)$.
%\begin{equation}
%\label{eq:impulse_response_defn}
%\impulseresponse_x(y) = \Aker(y, x),
%\end{equation} 
%which results from keeping $x$ fixed and viewing $\Aker(y,x)$ as a function of $y$.

\subsection{Required conditions}
\label{sec:conditions_2}

We focus on approximating operators that satisfy the following conditions: 
\begin{enumerate}
	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well defined for all $x\in \Omega$.
	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness. The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
\end{enumerate}
Our method may still perform well if these conditions are relaxed slightly. It is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, as long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region. The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small. If these conditions are violated, our method will incur additional error. If these conditions are violated too much, our method may fail. 
%Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.

\subsection{Finite element discretization}
\label{sec:finite_element_kernel}

In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces. In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below). With minor modifications, our method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $\fepoint_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(\fepoint_i)=1$ and $\febasis_i(\fepoint_j)=0$ if $i\neq j$. 
For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
$
%\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
$
Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries $\boldsymbol{\rho}_i = \rho_h(\febasis_i)$ for $i=1,\dots,\nsamplepts$.
%\begin{equation*}
%\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,\nsamplepts. 
%\end{equation*}
Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx$ for $i,j=1,\dots,\fedim$.
%\begin{equation*}
%\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,\fedim,
%\end{equation*}
The space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{w})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{w}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$. Direct calculation shows that $\mathbb{R}^\fedim_\massmatrix$ and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.

After discretization, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ is replaced by an operator $A_h:V_h \rightarrow V_h'$, which becomes an operator
\begin{equation*}
\mathbf{A}:\mathbb{R}^\fedim_\massmatrix \rightarrow \mathbb{R}^\fedim_{\massmatrix^{-1}}
\end{equation*}
under the isomorphism discussed above. Our method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$. What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$. Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{w}\mapsto \mathbf{A}^T\mathbf{w}$, and computing these matrix-vector products is costly.
Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)^T$ is the length $\fedim$ unit vector with one in the $j$th coordinate and zeros elsewhere. But computing the matrix-vector product $\mathbf{e}_j \mapsto \mathbf{A}\mathbf{e}_j$ is costly, and therefore wasteful if we do not use other matrix entries in the $j$th column of $\mathbf{A}$. Hence, methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.

>>>>

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}

% SIAM Shared Information Template
% This is information that is shared between the main document and any
% supplement. If no supplement is required, then this information can
% be included directly in the main document.

% Packages and macros go here
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{bm}

\begin{document}
	\begin{tikzpicture}[baseline]%, trim axis right]
		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
		ymode=log,
		ymin=1e-11,
		ymax=2,
		xmax=1000,
		axis x line*=bottom,
		axis y line*=left,
		xtick={0, 200, 400, 600,800,1000},
		xlabel={CG iteration},
		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},
		title={CG convergence},
		legend style={at={(1.2,1)},anchor=north east,draw=none},
		%legend pos=north east,
		legend style={font=\small}
		]
			\addlegendentry{Reg}
			\addplot[dotted, line width=0.5mm] table {advDiff_cg_rand_errs_reg.dat};
			\addlegendentry{None}
			\addplot[dash dot, line width=0.5mm] table {advDiff_cg_rand_errs_none.dat};
			\addlegendentry{PSF (1)}
			\addplot[blue, line width=0.5mm] table {advDiff_cg_rand_errs_psf1.dat};
			\addlegendentry{PSF (5)}
			\addplot[red, line width=0.5mm] table {advDiff_cg_rand_errs_psf5.dat};
			\addlegendentry{PSF (25)}
			\addplot[purple, line width=0.5mm] table {advDiff_cg_rand_errs_psf25.dat};
		\end{axis}
	\end{tikzpicture}
\end{document}
